Emittance preservation of an electron beam in a loaded quasi-linear plasma wakefield

### 2017-10-05 Tex/PaperPM.tex
Additional files from Erik and Patric

\begin{abstract}
We investigate beam loading and emittance preservation for a high-charge electron beam being accelerated in quasi-linear
plasma wakefield driven by a short proton beam. The structure of the studied wakefields are similar to those of a long, modulated
proton beam, such as the AWAKE proton driver.  We show that by %
properly choosing the electron beam parameters and exploiting two well known effects, beam loading of the wake field and full blow out
of plasma electrons by the accelerated beam, the electron beam can %PM be accelerated - EA ok
gain large amounts of energy with a narrow final energy spread (\%-level) and without significant emittance growth.
\end{abstract}

@@

Beam driven plasma wakefield accelerators \cite{chen:1985} have the potential to offer compact linear accelerators with high energy
gradients, and have been of interest for several decades. With a relativistic charged particle drive
beam travelling through a plasma, a strong wakefield is excited, that can be loaded by a trailing witness beam. When the
witness beam optimally loads the wakefield, an increase in absolute energy spread can be kept small. %to a minimum. %PM - EA ok
The concept
has been demonstrated experimentally in the past using electron drive beams accelerating electron witness beams
\cite{rosenzweig:1988, blumenfeld:2007, kallos:2008} [EA: M. Litos et al., Nature 515, 92?95 (2014)].

A major challenge with plasma wakefield accelerators is, however, to accelerate a beam while keeping both energy spread and
emittance growth small. %growth to a minimum. PM - EA ok
In the well described linear regime, valid when the beam density $n_{b}$ is much smaller than the
plasma density $n_{0}$, a non-linear transverse focusing force causes emittance growth of the witness beam. The beam
%will %PM - EA ok
also sees a transversely and longitudinally varying accelerating field causing a spread in energy after the beam has
been accelerated \cite{katsouleas:1987}. In the non-linear regime, where $n_{b} > n_{0}$, a bubble is formed by the
transverse oscillations of the plasma electrons, %forming PM - EA ok
gathering in a sheet around an evacuated area filled with only ions. The
ions, assumed stationary, form an uniform density ion channel creating a focusing force that varies linearly with
radius. This %produces a  PM - EA ok
focusing force %that PM - EA ok
preserves emittance \cite{rosenzweig:1991}.
% EA: I put Rosenzweig original blow out paper as citation here instead of, \cite{lu:2006-1, lu:2006}.

In this paper we present simulation results showing how emittance preservation of a high charge density witness beam can be ensured, when accelerated by a proton drive beam producing quasi-linear wakefields \cite{rosenzweig:2010}.  By quasi-linear wakefields we here mean wakefields with only partial blow out of the plasma electrons in the bubble.
%~ Patric's version is implemented:
The key idea is to have enough charge in the witness beam to at the same time load the wakefield to produce low
relative energy spread%$\Delta E/E$, 
, and completely blow out the electrons left in the bubble after the beam to reach conditions that preserves emittance.
The results have importance for the preparation of AWAKE Run 2 and possibly other applications in the quasi-linear regime.

@@

The energy carried by electron drive bunches used in previous plasma wakefield experiments have been no more than
$100\unit{J}$ and the propagation length typically no more than $1\unit{m}$ \cite{blumenfeld:2007,caldwell:2009}.  For high-energy physics application a higher total beam energy is often desired, for instance, the energy of a high-charge electron bunch accelerated to $1\unit{TeV}$ with $1\nexp{10}$ electrons, similar to the beam that could be produced by the
international linear collider, is $1.6\unit{kJ}$.  Using electron beams 
as drivers a large number of plasma stages is required to reach %PM - EA ok
an energy of a $\unit{kJ}$ for the accelerated bunch. However, staging plasma accelerators without reducing the
effective gradient and spoiling the beam quality is challenging \cite{steinke:2016,lindstrom:2016}.

Proton beams available at CERN carry significantly more energy than available electron beams, $19\unit{kJ}$ for the SPS beam \cite{gschwendtner:2016},
allowing for much longer plasma wakefield accelerator stages. The SPS beam is orders of magnitude longer
than the plasma wavelengths needed for such applications, and the beam %will PM - EA ok
does not drive a strong wake \cite{gschwendtner:2016}.  By letting the proton beam undergo
self-modulation before injecting the witness beam into accelerating structure, %most PM - EA ok
stronger wakefields are %formed PM - EA ok
excited. The self-modulation of the beam is
produced by the transverse fields generated by the beam acting upon itself, causing regions of the beam to rapidly
defocus \cite{kumar:2010}. The modulation frequency is close to that of the plasma, and produces a train of short proton
bunches along the beam axis with a surrounding halo of defocused particles. This train of bunches resonantly %produces - EA ok
drives
wakefields to large %r PM - EA ok 
amplitudes.

AWAKE at CERN is a proof of concept proton beam driven plasma wakefield
accelerator experiment \cite{awake_collaboration:2014}, currently in its first phase of operation. The experiment uses a $400\unit{GeV}$ proton beam delivered by the
SPS as its driver, and a single $10\unit{m}$ plasma stage with a nominal plasma density of $7\nexp{14}\unit{cm}^{-3}$
\cite{gschwendtner:2016}. This plasma density corresponds to $\lambda_{pe} = 1.26\unit{mm}$ -- matched to the transverse
size of the SPS proton beam such that $k_{pe}\sigma_{x,y} = 1$, where $k_{pe} = 2\pi/\lambda_{pe}$ is the plasma wave
number, %PM. - EA ok
$\lambda_{pe}=2\pi c/\omega_{pe}$ and $\omega_{pe}=\left(n_0e^2/m_e\epsilon_0\right)^{1/2}$ is the plasma electron angular frequency. %

@@

The focus in this study is the beam loading of the wakefields driven by the proton beam.   Studies of self-modulated proton bunches show that the proton bunch evolve as it propagates through a uniform plasma [EA: K. V. Lotov, Phys. Plas. 18, 024501 (2011], but that small variations in the plasma profile the modulation, and thus the wake fields, may be stabilized over long distances [EA: K. V. Lotov, Phys. Plas. 18, 024501 (2011), K. V. Lotov, Phys. Plas. 22, 103110 (2015), A. Caldwell and K. Lotov, Phys. Plas. 5, 103101 (2011)].  To study the witness beam evolution in a stable wake, independent of the dynamics of the self-modulation, %PMwe define here - EA ok
here we use a single, non-evolving proton bunch as driver.  The proton bunch parameters are chosen so that key features of the wake - the plasma electron density in the wake and the longitudinal electric field - are the same as %PM - EA ok
in the wake of a self-modulated proton beam with AWAKE baseline parameters \cite{gschwendtner:2016}.  Both the proton beam and the witness have Gaussian %PM - EA ok
longitudinal charge distribution and bi-Gaussian transverse charge distributions.

We have previously studied the beam loading in a proton beam wake using the full PIC code Osiris \cite{fonseca:2002}
%using PM - EA ok
and with 2D cylindrical-symmetric simulations.  The studies \cite{berglyd_olsen:2015, berglyd_olsen:2016} primarily looked at beam loading, energy gain and energy spread, as well as different approaches to create a stable drive beam structure based on previous self-modulation studies.  In order to study the witness beam emittance evolution we adopt the recently released open source version of QuickPIC \cite{huang:2006, an:2013}. QuickPIC is a fully relativistic 3D quasi-static PIC code. It does not suffer from
the numerical Cherenkov effect that full PIC codes do \cite{godfrey:1974,lehe:2013}, making it a well suited tool to
study emittance preservation.  All simulation results in this paper are %PMcalculated - EA ok
obtained using QuickPIC open source.

@@

%PM The baseline AWAKE drive beam current is insufficient to reach the non-linear regime and produce a plasma bubble. - EA ok
The modulation process used in AWAKE does not reach the fully non-linear regime and thus does not produce a bubble void of plasma electrons. When
the SPS beam, containing $3\nexp{11}$ protons \cite{gschwendtner:2016}, enters the second plasma stage (figure \ref{Fig:AWAKER2}), the peak
electric field is expected to be about $500\funit{MV}{m}$. The plasma electron density are only depleted to around $65\%$ of
nominal plasma density at the point where we inject the electron beam \cite{awake_collaboration:2016}.  The plasma electron density depletion and the peak field is replicated using a single %PM proton 
bunch with $1.46\nexp{10}$ protons (or $2.34\unit{nC}$), a %PM bunch
length $\sigma_{z} = 40\unit{\mu m}$ ($7\unit{kA}$), and a transverse size $\sigma_{x,y,proton} = 200\unit{\mu m}$.   The peak density is $0.83\cdot n_{0}$ results in a quasi-linear wake. %PM The beam energy is $400\unit{GeV}$.  
To avoid transverse evolution of the proton driver, emulating the stable propagation of the self-modulated proton beam [EA: K. V. Lotov, Phys. Plas. 18, 024501 (2011), K. V. Lotov, Phys. Plas. 22, 103110 (2015), A. Caldwell and K. Lotov, Phys. Plas. 5, 103101 (2011)] we freeze the transverse evolution of the equivalent proton bunch by increasing the particles mass by six orders of magnitude.

@@

%The small size of the accelerating structure produced by the self-modulated beam also
%puts constraints on the size of the electron witness beam, and thus its charge and current, if at the same time we wish
%to prevent large energy spread.  
%PM not sure what the point of the above sentence is ... - EA ok
In order to prevent large amplitude oscillations %PM
of the witness beam particles, which may cause additional energy spread as well as emittance growth, we %PMstudy
consider a witness beam matched to the plasma density.  
%The beam matching relation is given by
%\begin{equation}
%    \beta = \frac{\sigma_{x,y}^{2}\gamma}{\epsilon_{N}}
%          = \frac{c}{\omega_{pe}}\sqrt{2\gamma}, \label{EQ:MatchedB}
%    % = c\sqrt{\frac{m_{e}\varepsilon_{0}}{n_{pe}e^{2}}2\gamma}
%\end{equation}
%where $\beta$ is the Twiss parameter. It then follows that the transverse profile of the beam must satisfy
%PM Not sure what the point of the above lines is ....
The matched beam transverse size is:
\begin{equation}
%    \sigma_{x,y}^{2} = c\epsilon_{N}\sqrt{2\frac{m_{e}\varepsilon_{0}}{n_{pe}e^{2}\gamma}} \label{EQ:MatchedR}
%    \sigma_{x,y} =\left(c^2\epsilon_{N}^2m_{e}\varepsilon_{0}n_{pe}e^{2}\gamma \right)^{1/4} \label{EQ:MatchedR} - % PM - EA: not correct
	\sigma_{x,y}=\left(\frac{2c^{2}\varepsilon_{\mathrm{\mathrm{N}}}^{2}m_{e}\varepsilon_{0}}{n_{pe}e^{2}\gamma}\right)^{1/4}  \label{EQ:MatchedR}  % EA 
\end{equation}

We %furthermore PM - EA ok
assume an initial normalized emittance of $\epsilon_{N} = 2\unit{\mu m}$. This emittance is possible to produce with a standard RF-injector, while at the same time yielding a sufficient narrow beam. % EA: comment, perhaps put a ref here [***ref].

@@

Eq. (\ref{EQ:MatchedR}) yields a transverse size of $\sigma_{x,y}$ of $5.25\unit{\mu m}$, narrow compared to the drive beam $\sigma_{x,y,proton} = 200\unit{\mu m}$.   The bunch length was set to $\sigma_{z}=60\unit{\mu m}$, based on earlier %PM
beam loading studies \cite{berglyd_olsen:2016}.  The charge %PM was
is adjusted to $100\unit{pC}$ for optimal beam loading, as discussed in the next section. We refer to the defined drive beam and witness beam parameter set as the \emph{base case}.  
% EA: the key figure, figure 2, was not discussed in the text until very far down!  I therefore added (text could possibly be improved)
Figure \ref{Fig:PlasmaDenTWake} shows the two beams from QuickPIC simulations of the initial time step, for the base case parameters; the proton bunch in blue, the trailing electron beam in red and the plasma electron density in grey.

@@

The relatively small size of the witness beam %PM compared to the proton beam put some restrictions on the transverse resolution of the witness beam as we need a simulation box wide enough to contain the larger beam while resolving the narrower one (see figure \ref{Fig:PlasmaDenTWake}). 
puts constraints on the transverse cell size and number in the simulations: small size to resolve the small electron beam and large number of to resolve the large proton beam and the wakefields. %
We use %PMd  - EA ok
a transverse grid cell size of $1.17\unit{\mu m}$, and of
$2.34\unit{\mu m}$ for the longitudinal grid cells for the simulations presented in section \ref{S:BL}. The witness beam
was simulated with $16.8\nexp{6}$ non-weighted particles.  ***NN*** plasma particles were simulated per grid cell.
%   EA: Veronica to fill in NN above

\begin{figure}[hbt]
    \includegraphics[width=\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamLoading}
    \caption{\label{Fig:BeamLoading} Top plot%PM plot shows the - EA ok
: unloaded longitudinal electric field with no witness beam
        (dashed blue line) and %PMthe 
        loaded field (whole blue line) along the beam axis. The beam density along the axis
        for both beams are shown in red. %PM The bottom plot compares the 
        Bottom plot: plasma densities along the beam axis for a drive
        beam with no witness beam (dashed green line), witness beam with no drive beam (dash-dotted green line), and
        both beams present (whole green line). $\xi = z - tc$ is the position in the simulation box, moving towards the
        left.}
\end{figure}

@@

%~ Our configuration also shows some robustness to small electron beam transverse offsets as the bubble generated by the
%~ witness beam wakefield follows the beam itself.
%PM
So far we have considered a witness beam injected on the axis of the proton bunch. %
We now briefly examine the case of injection of a witness beam with an offset with respect to the proton bunch axis. %
Since the witness beam creates its own plasma bubble, the emittance of the part of the beam inside the
bubble is not affected by small transverse offsets of the witness beam with respect to the proton beam axis. This is illustrated in figure \ref{Fig:BeamEmitt}, right, for an electron beam offset of $1\sigma_{xe}$. Emittance preservation for small offsets is an added benefit of this 
accelerating regime, and may ease transverse injection tolerances. 
The head of the beam experiences a 
%larger initial emittance growth (see figure \ref{Fig:BeamFilament}a-f),
%PM
%THERE IS NO EMITTANCE SHOWN OR NOTHING TO COMPARE TO ... WHAT TO DO????
% EA: one way out, can compare emittance growth using fig 5 instead, as follows (alternatively, a new set of figures must be created)
larger initial emittance growth than for the on-axis case (compare figure \ref{Fig:BeamEmitt}, left, to \ref{Fig:BeamEmitt}, right), however, also for the head the emittance growth ends after the first few meters. Figure \ref{Fig:BeamFilament} a-c show the phase space of the head of the electron beam after 0, 1.0 and 2.5 m, while figure \ref{Fig:BeamFilament} d-f show the phase space of the trailing part of the beam.  The centroid oscillations of the head and the trailing part are shown in figure \ref{Fig:BeamFilament} g.  This effect of a transverse offset is likely to be greater for larger offsets as the beam oscillates
around the axis of the drive beam wakefield.% EA: would be good to include a statement of how large the offsets can be before when the electron beam starts to break up / not preserve emittance.

The transverse beam size within the bubble, where normalized emittance is preserved, follows the evolution given by
equation \ref{EQ:MatchedR} %PM where the plasma density is given by the ion density - EA ok
,i .e., evolves to stay matched. The on-axis density of the electron
beam, as a result, increases as its gamma factor increases and its transverse size decreases. The effect can be seen in
figure \ref{Fig:BeamFilament} h. This has the potential to cause overloading of the field. However, for the base case
no significant overloading %PMwas
is observed.

@@

Figure \ref{Fig:BeamQ}  show the dependence of %PM emittance
charge and energy spread on bunch length and %PMbunch
incoming charge.  An initial beam
emittance of $2\unit{\mu m}$ was used.  %
%PM
Therefore, we define the fractional charge as that whose emittance remains smaller than $2.1\unit{\mu m}$. %PM We ran the scan with beam charges from 
The beam charge range $10\unit{pC}$ to $300\unit{pC}$ and %with length
the $\sigma_{z}$ range from $40\unit{\mu m}$ to $100\unit{\mu m}$.  As can be seen from figure \ref{Fig:BeamQ} %, red curves
(red curves), both the
$40\unit{\mu m}$ and the $60\unit{\mu m}$ beam %PMhas
have a well defined minimum energy spread with a beam charge
$\approx 50\unit{pC}$ and $\approx 100\unit{pC}$, respectively. Lower beam charges tend to under-load the electric field,
while higher beam charges tend to overload it. It is also clear that longer beams with respect to the accelerating phase
of the field, $\approx\lambda_{pe}/4$, %PMwill
does not optimally load the wake, thus producing a larger spread in energy.  The blue curves show the fraction of %PMthe sliced emittance where the
charge whose initial emittance is preserved (the slice%Md
  emittance for the base case is shown in figure \ref{Fig:BeamEmitt}).  As the witness beam charge increases the fraction of slices with preserved emittance increases, as expected from an earlier onset of the bubble formation and also increase in the bubble size \cite{lu:2006-1, lu:2006}.
%PM SHOULD WE ADD SOMETHING LIKE:
We note here that operation with $100\unit{pC}$ leads to a significantly larger charge (factor $\sim 4$) with emittance preserved, at the expense of an increase of relative energy spread by a factor of two. %???

@@

% EA: slight rewrite to argue what we want with this section
The optimal working point will depend on the application and must be studied for each case and is, as illustrated in this section, a trade off between bunch length, bunch charge and emittance preservation, as well as other parameters like the plasma density.

@@

%This study assumes a simplified case with respect to a self-modulated proton beam, where the wake is driven by a single, short proton bunch, and investigates an ideal, uniform plasma profile.  To fully investigate optimal AWAKE injection, further study should be performed including simulations of realistic plasma profiles combined with a self-modulated proton beam.  
% EA: we could also say "is underway" and cite Lotov et co?
%PM: NOT SURE WHAT PLASMA PROFILE WEA RE TALKING ABOUT ... I would suggest - EA patric's verson of conclusions are OK with me

@@

Parameter studies indicate that up to a few $100\unit{pC}$, up to 70\% of the incoming bunch charge, is accelerated, for bunch of lengths of $40-60\unit{\mu m}$. Such electron bunches may be generated by an injector based on a standard RF photo-emission gun [EA: S. Doebert, private communication].

@@

% EA: I think owned also by IST.  I will check with Weiming and Warren
The simulations for this study have been performed using the open source version of QuickPIC released in early 2017 and owned
by UCLA.

>>>>

\begin{abstract}
We investigate beam loading and emittance preservation for a high-charge electron beam being accelerated in quasi-linear
plasma wakefield driven by a short proton beam. The structure of the studied wakefields are similar to those of a long, modulated
proton beam, such as the AWAKE proton driver.  We show that by %
properly choosing the electron beam parameters and exploiting two well known effects, beam loading of the wake field and full blow out
of plasma electrons by the accelerated beam, the electron beam can %PM be accelerated
gained large amounts of energy, have a narrow final energy spread (\%-level) and without significant emittance growth.
\end{abstract}

@@

Beam driven plasma wakefield accelerators \cite{chen:1985} have the potential to offer compact linear accelerators with high energy
gradients, and have been of interest for several decades. With a relativistic charged particle drive
beam travelling through a plasma, a strong wakefield is excited, that can be loaded by a trailing witness beam. When the
witness beam optimally loads the wakefield, an increase in absolute energy spread can be kept small. %to a minimum. %PM
The concept
has been demonstrated experimentally in the past using electron drive beams accelerating electron witness beams
\cite{rosenzweig:1988, blumenfeld:2007, kallos:2008} ***add [M. Litos and 23 co-authors, Nature 515, 92?95 (2014)]***.

A major challenge with plasma wakefield accelerators is, however, to accelerate a beam while keeping both energy spread and
emittance small. %growth to a minimum. PM
In the well described linear regime, valid when the beam density $n_{b}$ is much smaller than the
plasma density $n_{0}$, a non-linear transverse focusing force causes emittance growth of the witness beam. The beam
%will PM
also sees a transversely and longitudinally varying accelerating field causing a spread in energy after the beam has
been accelerated \cite{katsouleas:1987}. In the non-linear regime, where $n_{b} > n_{0}$, a bubble is formed by the
transverse oscillations of the plasma electrons, %forming PM
gathering in a sheet around an evacuated area filled with only ions. The
ions, assumed stationary, form an uniform density ion channel creating a focusing force that varies linearly with
radius. This %produces a  PM
focusing force %that PM
preserves emittance \cite{rosenzweig:1991}.
% EA: I put Rosenzweig original blow out paper as citation here instead of, \cite{lu:2006-1, lu:2006}.

In this paper we present simulation results showing how emittance preservation of a high charge density witness beam can be ensured, when accelerated by a proton drive beam producing quasi-linear wakefields \cite{rosenzweig:2010}.  By quasi-linear wakefields we mean mean wakefields with only partial blow out of the plasma electrons in the bubble.
%~ Patric's version is implemented:
The key idea is to have enough charge in the witness beam to at the same time load the wakefield to produce low
relative energy spread
%$\Delta E/E$, 
and completely blow out the electrons left in the bubble after the beam to reach conditions that preserves emittance.
The results have importance for the preparation of AWAKE Run 2 and possibly other applications in the quasi-linear regime.

@@

The energy carried by electron drive bunches used in previous plasma wakefield experiments have been no more than
$100\unit{J}$ and the propagation length typically no more than $1\unit{m}$ \cite{blumenfeld:2007,caldwell:2009}.  For high-energy physics application a higher total beam energy is often desired, for instance, the energy of a high-charge electron bunch accelerated to $1\unit{TeV}$ with $1\nexp{10}$ electrons, similar to the beam that could be produced by the
international linear collider, is $1.6\unit{kJ}$. Using electron beams 
as drivers a large number of plasma stages is required to reach %PM
an energy of a $\unit{kJ}$ for the accelerated bunch. However, staging plasma accelerators without reducing the
effective gradient and spoiling the beam quality is challenging \cite{steinke:2016,lindstrom:2016}.

Proton beams available at CERN carry significantly more energy than available electron beams, $19\unit{kJ}$ for the SPS beam \cite{gschwendtner:2016},
allowing for much longer plasma wakefield accelerator stages. The SPS beam is orders of magnitude longer
than the plasma wavelengths needed for such applications, and the beam %will PM
does not drive a strong wake \cite{gschwendtner:2016}.  By letting the proton beam undergo
self-modulation before injecting the witness beam into accelerating structure, %most PM
stronger wakefields are %formed PM
excited. The self-modulation of the beam is
produced by the transverse fields generated by the beam acting upon itself, causing regions of the beam to rapidly
defocus \cite{kumar:2010}. The modulation frequency is close to that of the plasma, and produces a train of short proton
bunches along the beam axis with a surrounding halo of defocused particles. This train of bunches resonantly %produces
drives
wakefields to large %r PM 
amplitudes.

AWAKE at CERN is a proof of concept proton beam driven plasma wakefield
accelerator experiment \cite{awake_collaboration:2014}, currently in its first phase of operation. The experiment uses a $400\unit{GeV}$ proton beam delivered by the
SPS as its driver, and a single $10\unit{m}$ plasma stage with a nominal plasma density of $7\nexp{14}\unit{cm}^{-3}$
\cite{gschwendtner:2016}. This plasma density corresponds to $\lambda_{pe} = 1.26\unit{mm}$ -- matched to the transverse
size of the SPS proton beam such that $k_{pe}\sigma_{x,y} = 1$, where $k_{pe} = 2\pi/\lambda_{pe}$ is the plasma wave
number, %PM.
$\lanmda_{pe}=2\pi c/\omega_{pe}$ and $\omega_{pe}=\left(n_0e^2/m_e\epsilon_0\right)^{1/2}$ is the plasma electron angular frequency. %

@@

The focus in this study is the beam loading of the wakefields driven by the proton beam.   Studies of self-modulated proton bunches show that the proton bunch evolve as it propagates through a uniform plasma [K. V. Lotov, Phys. Plas. 18, 024501 (2011], but that small variations in the plasma profile the modulation, and thus the wake fields, may be stabilized over long distances [K. V. Lotov, Phys. Plas. 18, 024501 (2011), K. V. Lotov, Phys. Plas. 22, 103110 (2015), A. Caldwell and K. Lotov, Phys. Plas. 5, 103101 (2011)].   To study the witness beam evolution in a stable wake, independent of the dynamics of the self-modulation, %PMwe define here
here we use a single, non-evolving proton bunch as driver.  The proton bunch parameters are chosen so that key features of the wake - the plasma electron density in the wake and the longitudinal electric field - are the same as %PM
in the wake of a self-modulated proton beam with AWAKE baseline parameters \cite{gschwendtner:2016}.  Both the proton beam and the witness have Gaussian %PM
longitudinal charge distribution and bi-Gaussian transverse charge distributions.

We have previously studied the beam loading in a proton beam wake using the full PIC code Osiris \cite{fonseca:2002}
%using PM
and withs 2D cylindrical-symmetric simulations.  The studies \cite{berglyd_olsen:2015, berglyd_olsen:2016} primarily looked at beam loading, energy gain and energy spread, as well as different approaches to create a stable drive beam structure based on previous self-modulation studies.  In order to study the witness beam emittance evolution we adopt the recently released open source version of QuickPIC \cite{huang:2006, an:2013}. QuickPIC is a fully relativistic 3D quasi-static PIC code. It does not suffer from
the numerical Cherenkov effect that full PIC codes do \cite{godfrey:1974,lehe:2013}, making it a well suited tool to
study emittance preservation.  All simulation results in this paper are %PMcalculated
obtained using QuickPIC open source.

@@

%PM The baseline AWAKE drive beam current is insufficient to reach the non-linear regime and produce a plasma bubble. 
The modulation process used in AWAKE does not reach the fully non-linear regime and thus does not produce a bubble void of plasma electrons. When
the SPS beam, containing $3\nexp{11}$ protons \cite{gschwendtner:2016}, enters the second plasma stage (figure \ref{Fig:AWAKER2}), the peak
electric field is expected to be about $500\funit{MV}{m}$. The plasma electron density are only depleted to around $65\%$ of
nominal plasma density at the point where we inject the electron beam \cite{awake_collaboration:2016}.  The plasma electron density depletion and the peak field is replicated using a single %PM proton 
bunch with $1.46\nexp{10}$ protons (or $2.34\unit{nC}$), a %PM bunch
length $\sigma_{z} = 40\unit{\mu m}$ ($7\unit{kA}$), and a transverse size $\sigma_{x,y,proton} = 200\unit{\mu m}$.   The peak density is $0.83\cdot n_{0}$ results in a quasi-linear wake. %PM The beam energy is $400\unit{GeV}$.  
To avoid transverse evolution of the proton driver, emulating the stable propagation of the self-modulated proton beam [K. V. Lotov, Phys. Plas. 18, 024501 (2011), K. V. Lotov, Phys. Plas. 22, 103110 (2015), A. Caldwell and K. Lotov, Phys. Plas. 5, 103101 (2011)] we freeze the transverse evolution of the equivalent proton bunch by increasing the particles mass by six orders of magnitude.

@@

%The small size of the accelerating structure produced by the self-modulated beam also
puts constraints on the size of the electron witness beam, and thus its charge and current, if at the same time we wish
to prevent large energy spread.  
%PM not sure what the point of the above sentence is ...
In order to prevent large amplitude oscillations %PM
of the witness beam particles, which may cause additional energy spread as well as emittance growth, we %PMstudy
consider a witness beam matched to the plasma density.  
%The beam matching relation is given by
%\begin{equation}
%    \beta = \frac{\sigma_{x,y}^{2}\gamma}{\epsilon_{N}}
%          = \frac{c}{\omega_{pe}}\sqrt{2\gamma}, \label{EQ:MatchedB}
%    % = c\sqrt{\frac{m_{e}\varepsilon_{0}}{n_{pe}e^{2}}2\gamma}
%\end{equation}
%where $\beta$ is the Twiss parameter. It then follows that the transverse profile of the beam must satisfy
%PM Not sure what the point of the above lines is ....
The matched beam transverse size is:
\begin{equation}
%    \sigma_{x,y}^{2} = c\epsilon_{N}\sqrt{2\frac{m_{e}\varepsilon_{0}}{n_{pe}e^{2}\gamma}} \label{EQ:MatchedR}
    \sigma_{x,y} =\left(c^2\epsilon_{N}^2m_{e}\varepsilon_{0}n_{pe}e^{2}\gamma}\right)^{1/4} \label{EQ:MatchedR}
\end{equation}

We %furthermore PM
assume an initial normalized emittance of $\epsilon_{N} = 2\unit{\mu m}$. This emittance is possible to produce with a standard RF-injector, while at the same time yielding a sufficient narrow beam. % EA: comment, perhaps put a ref here [***ref].

@@

Eq. (\ref{EQ:MatchedR}) yields a transverse size of $\sigma_{x,y}$ of $5.25\unit{\mu m}$, narrow compared to the drive beam $\sigma_{x,y,proton} = 200\unit{\mu m}$.   The bunch length was set to $\sigma_{z}=60\unit{\mu m}$, based on earlier %PM
beam loading studies \cite{berglyd_olsen:2016}.  The charge %PM was
is adjusted to $100\unit{pC}$ for optimal beam loading, as discussed in the next section. We refer to the defined drive beam and witness beam parameter set as the {base case}.

@@

The relatively small size of the witness beam %PM compared to the proton beam put some restrictions on the transverse resolution of the witness beam as we need a simulation box wide enough to contain the larger beam while resolving the narrower one (see figure \ref{Fig:PlasmaDenTWake}). 
puts constraints on the transverse cell size and number in the simulations: small size to resolve the small electron beam and large number of to resolve the large proton eam and the wakefields. %
We use %PMd 
a transverse grid cell size of $1.17\unit{\mu m}$, and of
$2.34\unit{\mu m}$ for the longitudinal grid cells for the simulations presented in section \ref{S:BL}. The witness beam
was simulated with $16.8\nexp{6}$ non-weighted particles.  ***NN*** plasma particles were simulated per grid cell.
%   EA: Veronica to fill in NN above

\begin{figure}[hbt]
    \includegraphics[width=\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamLoading}
    \caption{\label{Fig:BeamLoading} The top%PM plot shows the 
:unloaded longitudinal electric field with no witness beam
        (dashed blue line) and %PMthe 
        loaded field (whole blue line) along the beam axis. The beam density along the axis
        for both beams are shown in red. %PM The bottom plot compares the 
        Bottom plot: plasma densities along the beam axis for a drive
        beam with no witness beam (dashed green line), witness beam with no drive beam (dash-dotted green line), and
        both beams present (whole green line). $\xi = z - tc$ is the position in the simulation box, moving towards the
        left.}
\end{figure}

@@

%~ Our configuration also shows some robustness to small electron beam transverse offsets as the bubble generated by the
%~ witness beam wakefield follows the beam itself.
%PM
So far we have considered a witness beam injected on the axis of the proton bunch. %
We now briefly examine the case of injection of a witness beam with an offset with respect to the axis of the proton bunch. %

Since the witness beam creates its own plasma bubble, the emittance of the part of the beam inside the
bubble is not affected by small transverse offsets of the witness beam with respect to the proton beam axis. This is an added benefit of this 
accelerating regime, and may ease transverse injection tolerances. Also in the offset case the emittance of the head of the beam still stabilizes after some time.
This is illustrated in figure \ref{Fig:BeamEmitt} for an electron beam offset of $1\sigma_{x}$. In this case there is a
larger initial emittance growth (see figure \ref{Fig:BeamFilament}a-f),
%PM
%THERE IS NO EMITTANCE SHOWN OR NOTHING TO COMPARE TO ... WHAT TO DO????
 but the emittance growth stops after the first
few meters (figure \ref{Fig:BeamEmitt}). This effect is likely to be greater for larger offsets as the beam oscillates
around the axis of the drive beam wakefield, as can bee seen %PMfrom 
in figures \ref{Fig:BeamFilament}%PMg-h.
g and h.

The transverse beam size within the bubble, where normalized emittance is preserved, follows the evolution given by
equation \ref{EQ:MatchedR}%PM where the plasma density is given by the ion density
,i .e., evolves to stay matched. The on-axis density of the electron
beam, as a result, increases as its gamma factor increases and its transverse size decreases. The effect can be seen in
figure \ref{Fig:BeamFilament} h. This has the potential to cause overloading of the field. However, for the base case
no significant overloading %PMwas
is observed.

@@

Figure \ref{Fig:BeamQ}  show the dependence of %PM emittance
charge and energy spread on bunch length and %PMbunch
incoming charge.  An initial beam
emittance of $2\unit{\mu m}$ was used.  %
%PM
Therefore, we define the fractional charge as that whose emittance remains smaller than $2.1\unit{\mu m}$. %PM We ran the scan with beam charges from 
The beam charge range $10\unit{pC}$ to $300\unit{pC}$ and %with length
the $\sigma_{z}$ range from $40\unit{\mu m}$ to $100\unit{\mu m}$.  As can be seen from figure \ref{Fig:BeamQ} %, red curves
(red curves), both the
$40\unit{\mu m}$ and the $60\unit{\mu m}$ beam %PMhas
have a well defined minimum energy spread with a beam charge
$\approx 50\unit{pC}$ and $\approx 100\unit{pC}$, respectively. Lower beam charges tend to under-load the electric field,
while higher beam charges tend to overload it. It is also clear that longer beams with respect to the accelerating phase
of the field, $\approx\lambda_{pe}/4$, %PMwill
does not optimally load the wake, thus producing a larger spread in energy.  The blue curves show the fraction of %PMthe sliced emittance where the
charge whose initial emittance is preserved (the slice%Md
  emittance for the base case is shown in figure \ref{Fig:BeamEmitt}).  As the witness beam charge increases the fraction of slices with preserved emittance increases, as expected from an earlier onset of the bubble formation and also increase in the bubble size \cite{lu:2006-1, lu:2006}.
%PM SHOULD WE ADD SOMETHING LIKE:
We note here that operation with 100\unit{pC}$ leads to a significantly larger charge (factor $\sim4$) with emittance preserved, at the expense of an increase of relative energy spread by a factor of two. %???

@@

The optimal working point will depend on the application and must be studied for each case, based on a trade off between bunch length, bunch charge and emittance preservation, taking into account also other parameters, like the plasma density.

@@

%This study assumes a simplified case with respect to a self-modulated proton beam, where the wake is driven by a single, short proton bunch, and investigates an ideal, uniform plasma profile.  To fully investigate optimal AWAKE injection, further study should be performed including simulations of realistic plasma profiles combined with a self-modulated proton beam.  
% EA: we could also say "is underway" and cite Lotov et co?
%PM: NOT SURE WHAT PLASMA PROFILE WEA RE TALKING ABOUT ... I would suggest

@@

Parameter studies indicate that up to a few $100\unit{pC}$, up to 70\% of the incoming bunch charge, is accelerated, for bunch of lengths of $40-60\unit{\mu m}$. Such electron bunches may be generated by an injector based on a standard RF photo-emission gun [S. Doebert, private communication].

@@

The simulations for this study have been performed using the open source version of QuickPIC released in early 2017 and owned
by UCLA.
